---
title: "Linguistic prescriptivism, defectiveness and overabundance in French and Estonian"
output: html_document
date: "2023-08-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Prepping the data
Packages
```{r}
library(tidyverse)
library(ggpirate)
library(reshape2)
library(corrplot)
library(brms)
```

Importing the data

```{r}
read.pcibex <- function(filepath, auto.colnames=TRUE, fun.col=function(col,cols){cols[cols==col]<-paste(col,"Ibex",sep=".");return(cols)}) {
  n.cols <- max(count.fields(filepath,sep=",",quote=NULL),na.rm=TRUE)
  if (auto.colnames){
    cols <- c()
    con <- file(filepath, "r")
    while ( TRUE ) {
      line <- readLines(con, n = 1, warn=FALSE)
      if ( length(line) == 0) {
        break
      }
      m <- regmatches(line,regexec("^# (\\d+)\\. (.+)\\.$",line))[[1]]
      if (length(m) == 3) {
        index <- as.numeric(m[2])
        value <- m[3]
        if (is.function(fun.col)){
         cols <- fun.col(value,cols)
        }
        cols[index] <- value
        if (index == n.cols){
          break
        }
      }
    }
    close(con)
    return(read.csv(filepath, comment.char="#", header=FALSE, col.names=cols))
  }
  else{
    return(read.csv(filepath, comment.char="#", header=FALSE, col.names=seq(1:n.cols)))
  }
}

french <- read.pcibex("/Users/mariacopot/Desktop/git_repos/overabundance_defectiveness/analysis/results_def_oa_french.csv")
estonian <- read.pcibex("/Users/mariacopot/Desktop/git_repos/overabundance_defectiveness/analysis/results_def_oa_estonian.csv")
estonian <- estonian[estonian$PROLIFIC_PID != "PROLIFIC_PID",]

```
Extract item condition and language before merging the dataframes
```{r}
french$language <- "French"
estonian$language <- "Estonian"

french <- french %>% mutate(item_cond = case_when(
  grepl("de", recording) ~ "defective",
  grepl("an", recording) ~ "anglicism",
  grepl("un", recording) ~ "uncertain",
  grepl("ag", recording) ~ "ungrammatical",
))

estonian <- estonian %>% mutate(item_cond = case_when(
  grepl("oa", recording) ~ "overabundant",
  grepl("an", recording) ~ "anglicism",
  grepl("ug", recording) ~ "ungrammatical",
  grepl("uc", recording) ~ "uncertain",
))

results <- rbind(estonian, french)
```
Paste word corresponding to recording
```{r}
#items in the sentence frames
items <- read.csv("/Users/mariacopot/Desktop/git_repos/overabundance_defectiveness/analysis/items.csv", sep = "\t")
items$recording = paste(items$language, items$type, items$no, sep = "_")
items$recording = paste0(items$recording, ".wav")

results$lexeme <-  items$verb[match(results$recording, items$recording)]

```
Calculate prescriptiveness score. No one ever chose not to answer, so proceeding as such
```{r}
prs <- results %>% filter(Label == "lang_questions") %>%
   filter(grepl("prescr",PennElementName))
prs$Value <- as.numeric(prs$Value)
ggplot(prs, aes(y = Value, x = PennElementName))+geom_violin()+facet_grid(~language)
prs_wide.est <- prs %>%  filter(language == "Estonian") %>%
  dcast(`MD5.hash.of.participant.s.IP.address`~ PennElementName, value.var = "Value") %>% select(-`MD5.hash.of.participant.s.IP.address`)
prs_wide.fr <- prs %>%  filter(language == "French") %>%
  dcast(`MD5.hash.of.participant.s.IP.address`~ PennElementName, value.var = "Value") %>% select(-`MD5.hash.of.participant.s.IP.address`)
prs.rcorr.est <- cor(as.matrix(prs_wide.est))
prs.rcorr.fr <- cor(as.matrix(prs_wide.fr))
corrplot(prs.rcorr.est, order = "AOE")
corrplot(prs.rcorr.fr, order = "AOE")

#create holistic prescriptivism score
prs_grouped_all <- prs %>% group_by(`MD5.hash.of.participant.s.IP.address`) %>% summarize(score = sum(Value))
results$prescr_general <- with(results, prs_grouped_all$score[match(`MD5.hash.of.participant.s.IP.address`, prs_grouped_all$`MD5.hash.of.participant.s.IP.address`)])

```
Find if any participants had a particularly high rate of saying that they knew pseudolexemes, and exclude them.
```{r}
lex_check <- results[(results$PennElementName == "lex_checklist"),]
lex_check$known <- ifelse(lex_check$Value == "checked", 0, 1)
#there is orthographic variation between the PCIbex lexeme checklist file and the items file - fixing them before matching
lexeme_stats <- read.csv("/Users/mariacopot/Desktop/git_repos/overabundance_defectiveness/analysis/lexeme_stats.tsv", sep = "\t")

lex_check <- lex_check %>% 
  mutate(Parameter = case_when(Parameter == "sheerima"~ "šeerima",
                               Parameter == "chatima"~ "tšättima",
                               Parameter == "cheatima" ~"tšiitima",
                               Parameter == "chillima" ~"tšillima",
                               .default = as.character(Parameter)))
lex_check$is_real <- ifelse(lex_check$Parameter %in% lexeme_stats$lexeme, 1, 0)

#percentage of pseudolexemes reported as known by participant

participants_attention <- lex_check %>% group_by(MD5.hash.of.participant.s.IP.address) %>%
  filter(is_real == 0) %>%
  summarise(false_positives = sum(known)/36)

summary(participants_attention$false_positives)
#choice of threshold of 15% false positives based on distribution
bad_participants <- participants_attention[participants_attention$false_positives > 0.15,]$MD5.hash.of.participant.s.IP.address

results <- results %>% filter(!MD5.hash.of.participant.s.IP.address %in% bad_participants)
```
Add word frequency data
```{r}
#frequency, local entropy, surprisal for each lexeme

lexeme_stats <- lexeme_stats %>% separate(perc_lessfreqtoken, c("freq_num", "freq_denom"), sep = "/")
lexeme_stats <- lexeme_stats %>% mutate(
  freq_num = as.numeric(freq_num),
  freq_denom = as.numeric(freq_denom)
)
lexeme_stats$perc_lessfreqtoken <- lexeme_stats$freq_num/lexeme_stats$freq_denom

results$freq <- lexeme_stats$frequency[match(results$lexeme, lexeme_stats$lexeme)]
results$freq_num <- lexeme_stats$freq_num[match(results$lexeme, lexeme_stats$lexeme)]
results$freq_denom <- lexeme_stats$freq_num[match(results$lexeme, lexeme_stats$lexeme)]
results$perc_lessfreqtoken <- lexeme_stats$perc_lessfreqtoken[match(results$lexeme, lexeme_stats$lexeme)]
```

Demographic information
```{r}
demo <- results %>% filter(PennElementName == "demo_html")
results$age <-  demo[demo$Parameter == "age",]$Value[match(results$MD5.hash.of.participant.s.IP.address, demo$MD5.hash.of.participant.s.IP.address)]
results$educ <-  demo[grepl("dipl", demo$Parameter),]$Value[match(results$MD5.hash.of.participant.s.IP.address, demo$MD5.hash.of.participant.s.IP.address)]
results$gender <-  demo[ demo$Parameter == "sex",]$Value[match(results$MD5.hash.of.participant.s.IP.address, demo$MD5.hash.of.participant.s.IP.address)]

```

Get df with only judgements
```{r}
judgs <- results %>% 
  filter(PennElementName == "slider_judg") %>%
  filter(grepl("wav", recording)) %>%
  filter(Parameter == "Choice") %>%
  mutate(Value = as.numeric(Value))
```

Remove items for which the participant doesn't know the lexeme
```{r}

lex_check$key <- paste(lex_check$`MD5.hash.of.participant.s.IP.address`, lex_check$Parameter, sep = "~") 
judgs$key <- paste(judgs$`MD5.hash.of.participant.s.IP.address`, judgs$lexeme, sep = "~") 
judgs$known <- with(judgs, lex_check$known[match(key,lex_check$key)])

judgs <- judgs %>% filter(known == 1)
```


##Descriptive statistics

Plot item, participant and language conditions
```{r}

ggplot(judgs, aes(y = Value, x = item_cond, color = item_cond)) + 
  geom_jitter(alpha = 0.2, width = 0.3) + 
  geom_violin(alpha = 0.5, scale = "width") +
  facet_grid(partic_condition~language, scales = "free_x")
```


For uncertains and OA/DEF, look at the relationship with frequency
```{r}
judgs$logfreq <- log(judgs$freq)
uncertains <- judgs %>% filter(item_cond %in% c("uncertain", "defective", "overabundant"))


ggplot(judgs, aes(x = logfreq, y = Value, color = item_cond)) + geom_point(alpha = 0.5) + 
  facet_grid(partic_condition~language, scales = "free_x") + geom_smooth(method = "lm")
```

And with prescriptivism score
```{r}

ggplot(judgs, aes(x = prescr_general, y = Value, color = item_cond)) + geom_point(alpha = 0.5) + 
  facet_grid(partic_condition~language, scales = "free_x") + geom_smooth(method = "lm")
````

For estonian overabundants, check correlation between their score and their frequency
* u-shaped distribution when ordered by the percentage of the less frequent token
* same when ordered by numerator frequency
*
```{r}
ggplot(judgs[judgs$item_cond == "overabundant",], aes(x = fct_reorder(lexeme, perc_lessfreqtoken), y = Value)) + 
  geom_jitter(alpha = 0.2, width = 0.3) +
  geom_violin(alpha = 0.5, scale = "width") +
  geom_boxplot(width = 0.1) +
  facet_wrap(~partic_condition, scales = "free_x", ncol = 1)
````


##Inferential statistics

Coding the variables for a single model for all data
```{r}
#for modeling, we can think of defective and overabundant as the same item condition: uncertain items upon which the language's prescriptive culture has something explicit to say

judgs_inf <- judgs %>% mutate(
  language = as.factor(language),
  Value = as.numeric(Value),
  partic_condition= as.factor(partic_condition),
  PROLIFIC_PID = as.character(PROLIFIC_PID),
  prescr_general = prescr_general-9,
  age = as.numeric(age),
  educ = ordered(as.factor(educ),levels = c("no_dipl", "high_dipl","bach_dipl", "master_dipl" )),
  logfreq = log(freq)
)


contrasts(judgs_inf$language) <- c(-0.5, 0.5)
contrasts(judgs_inf$partic_condition) <- c(-0.5, 0.5)

judgs_inf$item_cond <- as.character(judgs_inf$item_cond)
%judgs_inf[judgs_inf$item_cond %in% c("defective", "overabundant"),]$item_cond <- "culture_targeted"
judgs_inf$item_cond <- as.factor(judgs_inf$item_cond)
judgs_inf$item_cond <- relevel(judgs_inf$item_cond, "ungrammatical")

c<-contr.treatment(4)
my.coding<-matrix(rep(1/4, 12), ncol=3)
contrasts(judgs_inf$item_cond)<- c-my.coding

judgs_inf <- judgs_inf %>% mutate(
  prescr_general_sc = scale(prescr_general),
  age_sc = scale(age),
  logfreq_sc = scale(logfreq)
)


judgs_inf$Value = judgs_inf$Value/99
````
Model

```{r}
mod2 <- brm(Value ~ language*item_cond*partic_condition*logfreq_sc+(1|recording)+(1|PROLIFIC_PID), data = judgs_inf,
            family = zero_one_inflated_beta(),
            chains = 4,
            cores = 4,
            init = "0",
            backend = "cmdstanr")

save(mod2, file = "/Users/mariacopot/Desktop/model.RDS")

conditions <- data.frame(expand.grid(partic_condition = c("normative", "possible"), language log_freq_std = c(-1.5, 0, 1.5) ))
byfreq_plot <- plot(conditional_effects(mod_all, effects = "prescr_general_std:item_cond", conditions = conditions))[[1]]+
  facet_grid(log_freq_std~partic_condition)+
  ylab("Judgement score") + xlab("Prescriptiveness") +
    scale_fill_discrete(name = "Item condition")+
  scale_color_discrete(name = "Item condition")+
  theme_minimal(15)
byfreq_plot$facet$params$labeller <- as_labeller(
  c(`-1.5` = "Low freq", `0` = "Mid freq", `1.5`= "High freq", `normative` = "Normative", `possible` = "Possible"))

library(glmmTMB)
judgs_inf$Value_glmm <- ifelse(judgs_inf$Value == "1", "0.999", 
                               ifelse(judgs_inf$Value == "0", "0.001", judgs_inf$Value))

judgs2 <- judgs_inf %>% select(language, item_cond, partic_condition, logfreq_sc, recording, Value_glmm, prescr_general_sc)
judgs2$logfreq_sc <- judgs2$logfreq_sc[,1]
judgs2$prescr_general_sc <- judgs2$prescr_general_sc[,1]
judgs2$Value_glmm <- as.numeric(judgs2$Value_glmm)


a <- glmmTMB(Value_glmm ~ language*item_cond*partic_condition*logfreq_sc+(1|recording), data = judgs2, family = beta_family())

judgs2$predicted_a <-  predict(a,type = "response")
ggplot(judgs2, aes(x = logfreq_sc, y = predicted_a, color = item_cond)) + geom_point(alpha = 0.2) + 
  facet_grid(partic_condition~language, scales = "free_x") + geom_smooth(method = "glm")

b <- glmmTMB(Value_glmm ~ language*item_cond*partic_condition*prescr_general_sc+(1|recording), data = judgs2, family = beta_family())
judgs2$predicted_b <-  predict(b,type = "response")
ggplot(judgs2, aes(x = prescr_general_sc, y = predicted_b, color = item_cond)) + geom_point(alpha = 0.2) + 
  facet_grid(partic_condition~language, scales = "free_x") + geom_smooth(method = "glm")

````

Two different models, one per language

```{r}

judgs_inf <- judgs %>% mutate(
  language = as.factor(language),
  Value = as.numeric(Value),
  partic_condition= as.factor(partic_condition),
  PROLIFIC_PID = as.character(PROLIFIC_PID),
  prescr_general = prescr_general-9,
  age = as.numeric(age),
  educ = ordered(as.factor(educ),levels = c("no_dipl", "high_dipl","bach_dipl", "master_dipl" )),
  logfreq = log(freq)
)

contrasts(judgs_inf$partic_condition) <- c(-0.5, 0.5)

judgs_inf$item_cond <- as.character(judgs_inf$item_cond)

c<-contr.treatment(4)
my.coding<-matrix(rep(1/4, 12), ncol=3)

judgs_inf$Value = judgs_inf$Value/99


est <- judgs_inf %>% filter(language == "Estonian")
est$item_cond <- as.factor(est$item_cond)
est$item_cond <- relevel(est$item_cond, "ungrammatical")
contrasts(est$item_cond)<- c-my.coding
est <- est %>% mutate(
  prescr_general_sc = scale(prescr_general),
  age_sc = scale(age),
  logfreq_sc = scale(logfreq)
)

fr <- judgs_inf %>% filter(language == "French")
fr$item_cond <- as.factor(fr$item_cond)
fr$item_cond <- relevel(fr$item_cond, "ungrammatical")
contrasts(fr$item_cond)<- c-my.coding
fr <- fr %>% mutate(
  prescr_general_sc = scale(prescr_general),
  age_sc = scale(age),
  logfreq_sc = scale(logfreq)
)

est_mod <- brm(Value ~ prescr_general_sc*item_cond*partic_condition*logfreq_sc+(1|recording)+(1|PROLIFIC_PID), data = est,
            family = zero_one_inflated_beta(),
            chains = 4,
            cores = 4,
            init = "0",
            backend = "cmdstanr")

save(est_mod, file = "/Users/mariacopot/Desktop/model_est.RDS")

fr_mod <- brm(Value ~ prescr_general_sc*item_cond*partic_condition*logfreq_sc+(1|recording)+(1|PROLIFIC_PID), data = fr,
            family = zero_one_inflated_beta(),
            chains = 4,
            cores = 4,
            init = "0",
            backend = "cmdstanr")

save(fr_mod, file = "/Users/mariacopot/Desktop/model_fr.RDS")
````

Plots for estonian

```{r}

#by frequency - all factors
conditions <- data.frame(expand.grid(partic_condition = c("normative", "possible"), logfreq_sc= c(-1.5, 0, 1.5) ))
byfreq_plot <- plot(conditional_effects(est_mod, effects = "prescr_general_sc:item_cond", conditions = conditions))[[1]]+
  facet_grid(logfreq_sc~partic_condition)+
  ylab("Judgement score") + xlab("Prescriptiveness") +
    scale_fill_discrete(name = "Item condition")+
  scale_color_discrete(name = "Item condition")+
  theme_minimal(15)
byfreq_plot$facet$params$labeller <- as_labeller(
  c(`-1.5` = "Low freq", `0` = "Mid freq", `1.5`= "High freq", `normative` = "Normative", `possible` = "Possible"))

byfreq_plot

#by prescriptiveness - all factors
conditions <- data.frame(expand.grid(partic_condition = c("normative", "possible"), prescr_general_sc= c(-2, 0, 2) ))
byprescr_plot <- plot(conditional_effects(est_mod, effects = "logfreq_sc:item_cond", conditions = conditions))[[1]]+
  facet_grid(prescr_general_sc~partic_condition)+
  ylab("Judgement score") + xlab("Prescriptivism") +
    scale_fill_discrete(name = "Item condition")+
  scale_color_discrete(name = "Item condition")+
  theme_minimal(15)
byprescr_plot$facet$params$labeller <- as_labeller(
  c(`-2` = "Low prescr", `0` = "Mid prescr", `2`= "High prescr", `normative` = "Normative", `possible` = "Possible"))

byprescr_plot
````
Plots for French

```{r}


#by frequency - all factors
conditions <- data.frame(expand.grid(partic_condition = c("normative", "possible"), logfreq_sc= c(-1.5, 0, 1.5) ))
byfreq_plot <- plot(conditional_effects(fr_mod, effects = "prescr_general_sc:item_cond", conditions = conditions))[[1]]+
  facet_grid(logfreq_sc~partic_condition)+
  ylab("Judgement score") + xlab("Prescriptiveness") +
    scale_fill_discrete(name = "Item condition")+
  scale_color_discrete(name = "Item condition")+
  theme_minimal(15)
byfreq_plot$facet$params$labeller <- as_labeller(
  c(`-1.5` = "Low freq", `0` = "Mid freq", `1.5`= "High freq", `normative` = "Normative", `possible` = "Possible"))

byfreq_plot

#by prescriptiveness - all factors
conditions <- data.frame(expand.grid(partic_condition = c("normative", "possible"), prescr_general_sc= c(-2, 0, 2) ))
byprescr_plot <- plot(conditional_effects(fr_mod, effects = "logfreq_sc:item_cond", conditions = conditions))[[1]]+
  facet_grid(prescr_general_sc~partic_condition)+
  ylab("Judgement score") + xlab("Prescriptivism") +
    scale_fill_discrete(name = "Item condition")+
  scale_color_discrete(name = "Item condition")+
  theme_minimal(15)
byprescr_plot$facet$params$labeller <- as_labeller(
  c(`-2` = "Low prescr", `0` = "Mid prescr", `2`= "High prescr", `normative` = "Normative", `possible` = "Possible"))

byprescr_plot
`````

